---
title: "exercise-09"
author: "Anais"
date: "4/1/2022"
output: html_document
---


* Using the {tidyverse} read_csv() function, load the “Street_et_al_2017.csv” dataset from this URL as a “tibble” named d

```{r setup, include=TRUE}
library(broom)
library(tidyverse)
library(manipulate)
library(patchwork)
library(infer)


f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f, col_names = TRUE)
#view(d)

```

* Do a quick exploratory data analysis where you generate the five-number summary (median, minimum and maximum and 1st and 3rd quartile values), plus mean and standard deviation, for each quantitative variable

```{r}

library (skimr)

skim(d)

```

* From this dataset, plot brain size (ECV) as a function of social group size (Group_size), longevity (Longevity), juvenile period length (Weaning), and reproductive lifespan (Repro_lifespan)

```{r}
par(mfrow=c(2,2))
plot(data = d, ECV ~ Group_size)
plot(data = d, ECV ~ Longevity)
plot(data = d, ECV ~ Weaning)
plot(data = d, ECV ~Repro_lifespan)

```

* Derive by hand the ordinary least squares regression coefficients β1 and β0 for ECV as a function of social group size

```{r}
# removing rows with na

d<-d %>% filter(!is.na(ECV))%>%
  filter(!is.na(Group_size))
#view(d)
(beta1 <- cor(d$ECV, d$Group_size) * (sd(d$ECV) / sd(d$Group_size)))
(beta0 <- mean(d$ECV) - beta1 * mean(d$Group_size))

```

* Confirm that you get the same results using the lm() function

```{r}


# using the lm function
m <- lm(ECV ~ Group_size, data = d)
m
summary(m)

# we get the same results than calculating β1 (2.463) and β0(30.3565) by hand.

```

* Repeat the analysis above for three different major radiations of primates – “catarrhines,” “platyrrhines,” and “strepsirhines”) separately. These are stored in the variable Taxonomic_group.
* Do your regression coefficients differ among groups? How might you determine this?

```{r}

fitted_models = d %>% 
  filter(!is.na(ECV))%>%
  filter(!is.na(Group_size))%>%
  filter(Taxonomic_group!='Catarrhini+B231')%>%
  group_by(Taxonomic_group) %>% do(model = lm(ECV ~ Group_size, data =.))
fitted_models$model

# or using the broom package

#fitted_models %>% tidy(model)#this does not work here - different data frame
#fitted_models %>% glance(model)

#or full summary for each model
plyr::l_ply(fitted_models$model, summary, .print = TRUE)

# -> yes the coeficients differ among groups

```
* For your first regression of ECV on social group size, calculate the standard error for the slope coefficient, the 95% CI, and the p value associated with this coefficient by hand. Also extract this same information from the results of running the lm() function.

```{r}

#SE by hand

n <- nrow(d)
mean_x <- mean(d$Group_size)
y_pred = beta0 + beta1*d$Group_size
y_error = d$ECV - y_pred
std_err_b1 <- sqrt((sum(y_error^2))/((n-2)*sum((d$Group_size-mean_x)^2)))
std_err_b1

std_err_b0<-std_err_b1*sqrt(
  sum((d$Group_size)^2)/n)
std_err_b0

#t statistic associated with B1… that will be the estimate of the coefficient divided by the standard error of the estimate in the coefficient: B1/(SE of B1)

t_stat<-beta1/std_err_b1
t_stat

ci_t <- beta1 + c(-1, 1) * qt(1 - 0.05 / 2, df = n - 2) * std_err_b1
ci_t

#or 
CI <- confint(m, level = 1 - 0.05)
CI
```



* Use a permutation approach with 1000 permutations to generate a null sampling distribution for the slope coefficient. What is it that you need to permute? What is the p value associated with your original slope coefficient?

with loop or infer

```{r}

#using the infer package

permuted.slope <- d %>%
  specify(ECV ~ Group_size) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

head(permuted.slope) # slopes from first few permutation replicates

hist(permuted.slope$stat, main="Histogram of Permuted\nSlope Values",
  xlab = "Slope Coefficient")



# original slope
original.slope <- lm(data = d, ECV ~ Group_size) %>%
  tidy(conf.int=TRUE, conf.level=.95) %>%
  filter(term=="Group_size")
original.slope # show model results for slope of weight


visualize(permuted.slope) +
  shade_p_value(obs_stat = original.slope$estimate, direction = "two_sided")



p.value <- permuted.slope %>% 
  mutate(abs_stat=abs(stat)) %>%
  summarize(
    estimate = mean(abs_stat >= abs(pull(original.slope, estimate)))
  )

p.value


```


* Use bootstrapping to generate a 95% CI for your estimate of the slope coefficient using both the percentile method and the theory-based method (i.e., based on the standard deviation of the bootstrapped sampling distribution). What is the p value associated with your observed slope coefficient based on each of these methods?


```{r}

#generate bootstrap replicate and calculate slope statistic

boot.slope <- d %>%
   specify(ECV ~ Group_size) %>%
   generate(reps = 1000, type = "bootstrap") %>%
   calculate(stat = "slope")
 
 head(boot.slope) 
 
#calculate CI with percentile and SE method
 
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha/2
p_upper <- 1 - (alpha/2)
degrees_of_freedom <- nrow(d) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)
 
 boot.slope.summary <- boot.slope %>%
  summarize(
    estimate = mean(stat),
    std.error = sd(stat),
    # calculate the CI based on the SE and t distribution
    lower = estimate - std.error * critical_value,
    upper = estimate + std.error * critical_value,
    # calculate the CI based on the quantile (percentile)  method
    boot.lower = quantile(stat, p_lower),
    boot.upper = quantile(stat, p_upper)
  )

# show summary of bootstrap sampling distribution
boot.slope.summary

# or using get_ci()

#percentile method
get_ci(boot.slope, level = 1 - alpha, type = "percentile")
#se method
get_ci(boot.slope, level = 1 - alpha, type = "se", point_estimate = pull(boot.slope.summary,
    estimate))


#p-value associated with the original slope
(p.value <- boot.slope %>%
  get_p_value(obs_stat = original.slope$estimate, direction="two_sided"))

# -> p-value=0.946 -> we dont reject the null hypothesis, there is no diference between the original and boostrap slope

visualize(boot.slope) +
  shade_p_value(obs_stat = original.slope$estimate, direction = "two_sided")

```






